Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-07-28 psy_peds.Rmd
correcting cohorts issue, and adding comparison plot psy/non-psy diagnoses

@@

### Extract the patient summary and observation information. 
demo_raw <- files[[1]]
obs_raw <- files[[2]]
clinical_raw <- files[[3]]

@@

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

@@

# cut-off dates used in the analysis 
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
#early_pandemic_start_date <- "2020-03-15"
#late_pandemic_start_date <- "2021-01-01"

input_data <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "days_since_admission")) %>%
  full_join( demo_raw, by = "patient_num") %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = cut( date, breaks= "week"),
                 year = format( date, "%Y")) %>% 
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 concept_code %in% icdCodes$ICD10_Code,
                 age < 18 & age > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  # dplyr::mutate(period = ifelse( date < early_pandemic_start_date,
  #                                 "before_pandemic",
  #                                 ifelse(date < late_pandemic_start_date,
  #                                        "early_pandemic", 
  #                                        "late_pandemic"))) %>%
  dplyr::mutate(period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","after_pandemic")) %>%
  left_join( icdCodes, by = c("concept_code" = "ICD10_Code")) %>%
  dplyr::select( patient_num, sex, age, concept_type, concept_code, calendar_date, in_hospital, n_hospitalisation, len_hospitalisation, 
                 date, weeks, year, period, disorder_group, description )
if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = "patient_num" )
}
```

@@

```{r}
patient_count_barplot_period <- input_data %>% 
  dplyr::group_by( disorder_group, patient_num ) %>% 
  dplyr::mutate( 
    when = ifelse( "before_pandemic" %in% period & "after_pandemic" %in% period,
                   "Before and After Pandemic",
                   ifelse( "before_pandemic" %in% period & ! "after_pandemic" %in% period,
                           "Only Before Pandemic",
                           "Only After Pandemic"
                   )
    ) 
    %>% factor(levels = c("Only Before Pandemic", "Before and After Pandemic", "Only After Pandemic"))
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::select( patient_num, disorder_group, when ) %>%
  dplyr::group_by( disorder_group, when ) %>%
  dplyr::summarise( count = n_distinct( patient_num ) ) %>% 
  dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
                               count, 
                               0.5), 
                siteid = site)

@@

## Diagnoses count per week
Here we calculate the number of patients hospitalized per week (exactly, the number of patients being assigned a diagnosis on that week, hence usually corresponding to the end of hospitalization, ie not taking into account the length of the encounter). 
The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
```{r}
patient_count_per_week <- input_data[, c("patient_num", "weeks")] %>%
  unique() %>%
  group_by(weeks) %>%
  tally() %>%
  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
                     n,
                     0.5),
         siteid = site )

#estimate the number of patients per week
patient_count_per_week %>%
  dplyr::mutate(weeks = as.Date(weeks)) %>% 
  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
                                  "before_pandemic",
                                  #ifelse(weeks < late_pandemic_start_date,
                                  #       "early_pandemic", 
                                  #       "late_pandemic")
                                 "after_pandemic")) %>% 
  ggplot(aes(x = weeks, y = n, fill = period)) +
  geom_bar(stat="identity")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Calendar Weeks",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

@@

@@

```{r, }
# Creating encounter table with observation assigned to encounters
# 1. Creating encounters identifiers based on clinical_raw information about start and end date of an encounter --> n_hospitalisation
# 2. Assigning ICD codes to encounters (because they are assigned to the first or last day but usually are meaningful for the entire encounter)
encounters_obs <- clinical_raw %>% 
  filter(in_hospital == 1) %>% 
  select(patient_num, days_since_admission, n_hospitalisation) %>% 
  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "days_since_admission")) %>% 
  drop_na(concept_type) %>% 
  distinct(patient_num, n_hospitalisation, concept_type, concept_code)

obs_by_day <- encounters_obs %>% 
  filter(concept_code %in% icdCodes$ICD10_Code) %>% 
  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "n_hospitalisation", "calendar_date")], 
            on = c("patient_num", "n_hospitalisation")) %>% 
  group_by(calendar_date) %>% 
  summarise(patient_count = n_distinct(patient_num)) %>% 
  mutate(calendar_date = as.Date(calendar_date, format=dateFormat), 
         period = if_else(calendar_date < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(calendar_date < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         ), 
         patient_count = ifelse(patient_count > obfuscation | isFALSE(obfuscation), 
                                 patient_count, 
                                 0.5)
  )

@@

### Patients count per week aggregated by sex
Same graphics as "Patient Count per Week", dichotomized by patients sex.
```{r}
patient_count_per_week_sex <- input_data[, c("patient_num", "sex","weeks")] %>%
  unique() %>%
  group_by(weeks, sex) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation),
                     n, 
                     0.5),
         siteid = site)

patient_count_per_week_sex %>%
  ggplot(aes(x = as.Date(weeks), y = n, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts with mental health related ICD codes aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

@@

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "age","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

@@

patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1,c("patient_num", "weeks", "disorder_group")] %>%
  unique() %>%
  group_by(weeks, disorder_group) %>% tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation), n, 0.5), siteid = site)

patient_count_per_week_disorder_group %>%
  mutate(weeks = as.Date(weeks), 
         period = if_else(weeks < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(weeks < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         )
  ) %>% 
  ggplot(aes(x = as.Date(weeks), y= n, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Week",
       y = "Patient Counts",
       title = "Weekly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
ggsave("output/categorical_counts_inpatient_teens_all.png", width = 18, height = 9, units="cm")
```

@@

## Age
```{r}
age_stats <- input_data %>%
  dplyr::distinct(patient_num, period, age) %>%
  filter(age != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age ), 
                    sd = sd( age ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
  )

@@

## Sex
```{r}
sex <- input_data %>% 
  dplyr::group_by( sex ) %>%
  dplyr::summarise( count = n_distinct(patient_num) ) %>%
  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                 total = sum( count ), percentage = round( count/total*100, 2)  )

sex_by_period <- input_data %>% 
  group_by(period, sex ) %>% 
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, .add=TRUE) %>% 
  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
          percentage=round(count/sum( count )*100, 2)) %>% 
  replace_na(list(male = 0, female = 0))

@@

## Race
```{r}
if( raceAvailable == TRUE ){
  race <- input_data %>%
    dplyr::group_by( race_4ce ) %>%
    dplyr::summarise( count = n_distinct(patient_num) ) %>%
    dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                   total = sum( count ), percentage = round( count/total*100, 2)  )

  race_by_period <- input_data %>%
    group_by(period, race_4ce ) %>% 
    summarise(count = n_distinct(patient_num)) %>%
    group_by(period, .add=TRUE) %>% 
    mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
            percentage=round(count/sum( count )*100, 2))

@@

### By ICD code
```{r}
summary_diagnosis <- input_data %>%
  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>% 
  dplyr::group_by( concept_code, description, period ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc(count) ) %>% 
  dplyr::ungroup()

@@

### By disorder group
```{r}
summary_disorder_group <- input_data %>%
  dplyr::group_by( disorder_group ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc( count ))

@@

### Length of hospitalization 
```{r}
## Calculating also mean and sd in case we can only retrieve aggregated data from the sites
hospit_length <- input_data %>%
  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
  dplyr::group_by( period ) %>% 
  dplyr::summarise( n_hospitalisations = n(),
                    length_hospt_median = median(len_hospitalisation), 
                    IQR = IQR( len_hospitalisation), 
                    length_hospt_mean = round(mean(len_hospitalisation), 2),
                    length_hospt_sd = round(sd(len_hospitalisation), 2))

@@

### Repeat hospitalizations by patient
```{r}
hospitalisation_count <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by( period, patient_num ) %>%
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>%
  dplyr::group_by( period, add = FALSE) %>% 
  dplyr::summarise( mean_count_hospitalizations = mean(n_hospitalisation), 
                    max_count_hospitalizations = max(n_hospitalisation),
                    median_count_hospitalizations = median(n_hospitalisation), 
                    sd_hospitalisation_count = sd(n_hospitalisation),
                    IQR_hospitalisation_count = IQR( n_hospitalisation ))

repeated_hospitalisation <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by(patient_num, period) %>% 
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>% 
  dplyr::mutate(repeated_hospitalisation = if_else(n_hospitalisation > 1, "repeated", "non_repeated")) %>% 
  dplyr::group_by( period, add = FALSE ) %>% 
  dplyr::count( repeated_hospitalisation ) %>% 
  tidyr::pivot_wider(names_from = repeated_hospitalisation, values_from = n) %>% 
  replace_na(list("non_repeated" = 0, "repeated" = 0))
repeated_hospitalisation_test <- chisq.test(repeated_hospitalisation[, c("non_repeated", "repeated")]) %>% 
  `[[`("p.value")
```

>>>>

```{r}
# cut-off dates used in the analysis, these should not be changed
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
```

@@

### Extract the patient summary and observation information. 
demo_raw <- files[["patientSummary"]]
obs_raw <- files[["patientObservations"]]
clinical_raw <- files[["patientClinicalCourse"]]

@@

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))

### Joining the icd code description to the table 
obs_raw <- obs_raw %>%
  left_join(icdCodes, by = c("concept_code" = "ICD10_Code"))
```

### Data management for the patientClinicalCourse table
```{r}
clinical_raw <- clinical_raw %>% 
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks= "week")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","after_pandemic")) %>% 
  dplyr::filter(cohort  %in% c("AllAdm", "AllNotAdm"))
```

@@

#early_pandemic_start_date <- "2020-03-15"
#late_pandemic_start_date <- "2021-01-01"

all_codes <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  full_join( demo_raw, by = c("patient_num", "cohort")) %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 age < 18 & age > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, year,
                 period, disorder_group, description )

input_data <- all_codes %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = c("patient_num", "cohort") )
}
```

@@

```{r}
patient_count_barplot_period <- input_data %>% 
  dplyr::group_by( disorder_group, patient_num, cohort) %>% 
  dplyr::mutate( 
    when = ifelse( "before_pandemic" %in% period & "after_pandemic" %in% period,
                   "Before and After Pandemic",
                   ifelse( "before_pandemic" %in% period & ! "after_pandemic" %in% period,
                           "Only Before Pandemic",
                           "Only After Pandemic"
                   )
    ) 
    %>% factor(levels = c("Only Before Pandemic", "Before and After Pandemic", "Only After Pandemic"))
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::select( patient_num, cohort, disorder_group, when ) %>%
  dplyr::group_by( disorder_group, when ) %>%
  dplyr::summarise( count = n_distinct( patient_num, cohort ) ) %>% 
  dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
                               count, 
                               0.5), 
                siteid = site)

@@

## Diagnoses count per week
Here we calculate the number of patients hospitalized per week (exactly, the number of patients being assigned a diagnosis on that week, hence usually corresponding to the end of hospitalization, ie not taking into account the length of the encounter). 
The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
```{r}
patient_count_per_week <- input_data[, c("patient_num", "cohort", "weeks")] %>%
  unique() %>%
  group_by(weeks) %>%
  tally() %>%
  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
                     n,
                     0.5),
         siteid = site )

#estimate the number of patients per week
patient_count_per_week %>%
  dplyr::mutate(weeks = as.Date(weeks)) %>% 
  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
                                  "before_pandemic",
                                 "after_pandemic")) %>% 
  ggplot(aes(x = weeks, y = n, fill = period)) +
  geom_bar(stat="identity")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Calendar Weeks",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

@@

```{r}
patient_count_per_week <- all_codes %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "somatic")
  ) %>% 
  select(patient_num, cohort, weeks, icd_code_category, period) %>% 
  unique() %>%
  group_by(weeks, icd_code_category, period) %>%
  tally() %>%
  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
                     n,
                     0.5),
         siteid = site )

#estimate the number of patients per week
patient_count_per_week %>%
  ggplot(aes(x = weeks, y = n, fill = icd_code_category, color = icd_code_category)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Calendar Weeks",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

ggsave("output/overall_weekly_count_all_comparison.png", width = 18, height = 9, units="cm")
```

@@

```{r, }
# Creating encounter table with observation assigned to encounters
# 1. Creating encounters identifiers based on clinical_raw information about start and end date of an encounter --> n_hospitalisation
# 2. Assigning ICD codes to encounters (because they are assigned to the first or last day but usually are meaningful for the entire encounter)
encounters_obs <- clinical_raw %>% 
  filter(in_hospital == 1) %>% 
  select(patient_num, cohort, days_since_admission, n_hospitalisation) %>% 
  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "cohort", "days_since_admission")) %>% 
  drop_na(concept_type) %>% 
  distinct(patient_num, cohort, n_hospitalisation, concept_type, concept_code)

obs_by_day <- encounters_obs %>% 
  filter(concept_code %in% icdCodes$ICD10_Code) %>% 
  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "cohort", "n_hospitalisation", "calendar_date")], 
            on = c("patient_num", "n_hospitalisation")) %>% 
  group_by(calendar_date) %>% 
  summarise(patient_count = n_distinct(patient_num, cohort)) %>% 
  mutate(calendar_date = as.Date(calendar_date, format=dateFormat), 
         period = if_else(calendar_date < pandemic_start_date,
                          "Before Pandemic",
                          "After Pandemic"
         ), 
         patient_count = ifelse(patient_count > obfuscation | isFALSE(obfuscation), 
                                 patient_count, 
                                 0.5)
  )

@@

### Patients count per week aggregated by sex
Same graphics as "Patient Count per Week", dichotomized by patients sex.
```{r}
patient_count_per_week_sex <- input_data[, c("patient_num", "cohort", "sex","weeks")] %>%
  unique() %>%
  group_by(weeks, sex) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation),
                     n, 
                     0.5),
         siteid = site)

patient_count_per_week_sex %>%
  ggplot(aes(x = as.Date(weeks), y = n, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts with mental health related ICD codes aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")

@@

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "cohort", "age","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
ggsave("output/overall_weekly_count_age.png", width = 18, height = 9, units="cm")

@@

patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1, c("patient_num", "cohort", "weeks", "disorder_group")] %>%
  unique() %>%
  group_by(weeks, disorder_group) %>% tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation), n, 0.5), siteid = site)

patient_count_per_week_disorder_group %>%
  mutate(weeks = as.Date(weeks), 
         period = if_else(weeks < pandemic_start_date,
                          "Before Pandemic",
                          "After Pandemic"
         )
  ) %>% 
  ggplot(aes(x = as.Date(weeks), y= n, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Week",
       y = "Patient Counts",
       title = "Weekly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
ggsave("output/categorical_counts_inpatient_teens_all.png", width = 18, height = 9, units="cm")
```

@@

## Age
```{r}
age_stats <- input_data %>%
  dplyr::distinct(patient_num, cohort, period, age) %>%
  filter(age != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age ), 
                    sd = sd( age ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
  )

@@

## Sex
```{r}
sex <- input_data %>% 
  dplyr::group_by( sex ) %>%
  dplyr::summarise( count = n_distinct(patient_num, cohort) ) %>%
  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                 total = sum( count ), percentage = round( count/total*100, 2)  )

sex_by_period <- input_data %>% 
  group_by(period, sex ) %>% 
  summarise(count = n_distinct(patient_num, cohort)) %>%
  group_by(period, .add=TRUE) %>% 
  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
          percentage=round(count/sum( count )*100, 2)) %>% 
  replace_na(list(male = 0, female = 0))

@@

## Race
```{r}
if( raceAvailable == TRUE ){
  race <- input_data %>%
    dplyr::group_by( race_4ce ) %>%
    dplyr::summarise( count = n_distinct(patient_num, cohort) ) %>%
    dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                   total = sum( count ), percentage = round( count/total*100, 2)  )

  race_by_period <- input_data %>%
    group_by(period, race_4ce ) %>% 
    summarise(count = n_distinct(patient_num, cohort)) %>%
    group_by(period, .add=TRUE) %>% 
    mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
            percentage=round(count/sum( count )*100, 2))

@@

### By ICD code
```{r}
summary_diagnosis <- input_data %>%
  dplyr::distinct( patient_num, cohort, description, period, weeks, concept_code ) %>% 
  dplyr::group_by( concept_code, description, period ) %>%
  dplyr::count( patient_num, cohort ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc(count) ) %>% 
  dplyr::ungroup()

@@

### By disorder group
```{r}
summary_disorder_group <- input_data %>%
  dplyr::group_by( disorder_group ) %>%
  dplyr::count( patient_num, cohort ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc( count ))

@@

### Length of hospitalization 
```{r}
## Calculating also mean and sd in case we can only retrieve aggregated data from the sites
hospit_length <- input_data %>%
  dplyr::distinct( patient_num, cohort, n_hospitalisation, len_hospitalisation, period ) %>%
  dplyr::group_by( period ) %>% 
  dplyr::summarise( n_hospitalisations = n(),
                    length_hospt_median = median(len_hospitalisation), 
                    IQR = IQR( len_hospitalisation), 
                    length_hospt_mean = round(mean(len_hospitalisation), 2),
                    length_hospt_sd = round(sd(len_hospitalisation), 2))

@@

### Repeat hospitalizations by patient
```{r}
hospitalisation_count <- input_data %>% 
  dplyr::distinct( patient_num, cohort, period, n_hospitalisation ) %>% 
  dplyr::group_by( period, patient_num, cohort ) %>%
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>%
  dplyr::group_by( period, add = FALSE) %>% 
  dplyr::summarise( mean_count_hospitalizations = mean(n_hospitalisation), 
                    max_count_hospitalizations = max(n_hospitalisation),
                    median_count_hospitalizations = median(n_hospitalisation), 
                    sd_hospitalisation_count = sd(n_hospitalisation),
                    IQR_hospitalisation_count = IQR( n_hospitalisation ))

repeated_hospitalisation <- input_data %>% 
  dplyr::distinct( patient_num, cohort, period, n_hospitalisation ) %>% 
  dplyr::group_by(patient_num, cohort, period) %>% 
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>% 
  dplyr::mutate(repeated_hospitalisation = if_else(n_hospitalisation > 1, "repeated", "non_repeated")) %>% 
  dplyr::group_by( period, add = FALSE ) %>% 
  dplyr::count( repeated_hospitalisation ) %>% 
  tidyr::pivot_wider(names_from = repeated_hospitalisation, values_from = n) %>% 
  replace_na(list("non_repeated" = 0, "repeated" = 0))
repeated_hospitalisation_test <- chisq.test(repeated_hospitalisation[, c("non_repeated", "repeated")]) %>% 
  `[[`("p.value")
```
